---
title: "Data on Posit Connect using {pins}"
---

This page goes through how to put data on [Posit Connect](https://posit.co/products/enterprise/connect/) using [pins](https://pins.rstudio.com/).

## Connecting to Posit Connect

For the smoothest experience, we recommend that you authenticate using environment variables. The two variables you will need are `CONNECT_SERVER` and `CONNECT_API_KEY`.

::: {.callout-tip}
The function [usethis::edit_r_environ()](https://usethis.r-lib.org/reference/edit.html) can be very handy to open `.Renviron` file to specify your environment variables.
:::

`CONNECT_SERVER` is the URL of the posit connect page. So if your connect server is accessed through `https://example.com/connect/#/content/` then you can find `CONNECT_SERVER` by removing `connect/` and everything that follows it, leaving you with `https://example.com/`.

`CONNECT_API_KEY` is created through your Connect server. 
1. Click on your name in the upper right upper right.
1. Click `API keys`.
1. Click `New API Key`.
1. Give your API a key, click ``Create Key`.

Once you have those two, you can add them to your `.Renviron` file in the following format:

```markdown
CONNECT_API_KEY=xxxxxxxxxxxxxxxxxxxxxxxxxxxx
CONNECT_SERVER=https://example.com/
```

Note that you don't want to put quotes around the values.

## Creating the board
```{r}
#| message: false
library(pins)

board <- board_connect()
```

We load in the data we created in [data-prep](data-prep.qmd)

```{r}
#| message: false
flights <- readr::read_csv("data/laxflights2022_lite.csv")
```

## Pinning data set to Connect

Now that we have the data set, we can send it to Posit Connect using the `board` we created, the data set and a name of out choosing.

```{r}
#| eval: false
#| message: false
board |>
  pin_write(flights, name = "laxflights2022_lite", type = "rds")
```
