---
title: "Put data places with {pins}"
---

This page shows how to use [pins](https://pins.rstudio.com/) to store data in various platforms as needed for the [pipelines](pipelines.qmd).

## Posit Connect

This section goes through how to put data on [Posit Connect](https://posit.co/products/enterprise/connect/) using [pins](https://pins.rstudio.com/).

### Connecting to Posit Connect

{{< include _include-envvar-connect.qmd >}}

### Creating the board

```{r}
#| message: false
library(pins)

board <- board_connect()
```

We load in the data we created in [data-prep](data-prep.qmd)

```{r}
#| message: false
flights <- readr::read_csv("data/laxflights2022_lite.csv")
```

### Pinning data set to Connect

Now that we have the data set, we can send it to Posit Connect using the `board` we created, the data set and a name of out choosing.

```{r}
#| eval: false
#| message: false
board |>
  pin_write(flights, name = "laxflights2022_lite", type = "rds")
```


## Amazon S3

This section goes through how to put data on [Amazon S3](https://aws.amazon.com/s3/) using [pins](https://pins.rstudio.com/).

### Connecting to Amazon S3

{{< include _include-envvar-s3.qmd >}}

### Creating the board

```{r}
#| message: false
library(pins)

board <- board_s3(
  "tidymodels-pipeline-example",
  region = "us-west-1"
)
```

We load in the data we created in [data-prep](data-prep.qmd)

```{r}
#| message: false
flights <- readr::read_csv("data/laxflights2022_lite.csv")
```

### Pinning data set to Amazon S3

Now that we have the data set, we can send it to Amazon S3 using the `board` we created, the data set and a name of out choosing.

```{r}
#| eval: false
#| message: false
board |>
  pin_write(flights, name = "laxflights2022_lite", type = "rds")
```
